\newpage
\section{Control systems}

\subsection{Introduction}
Oxygenation
Our process revolves around the growth and harvesting of bovine cells. Preventing cell death and enabling growth in mammalian cells means we need to ensure they’re able to respire. The body does this with the use of the bloodstream – oxygen enters the lungs and binds to the haemoglobin on red blood cells, which are then pumped around the body through a network of vessels to oxygenate our cells. This prevents necrosis and allows cells access to energy via their mitochondria for work and division.
Since we’re using a [batch-fed] bioreactor, we can use the culture media as our vehicle for oxygen, as we know all live cells have access to the medium thanks to cell requirements. We also need to control the CO2 content of the medium at a reasonable, [and our gas cycling could be just the way to extract it.]
Medium Oxygenation Methods
Purified Oxygen
Oxygenation can be done with air or oxygen. 
If using purified oxygen, much less gas is needed but it’d have to be purchased or refined on site – which increases operating & capital cost, complexity, and maintenance. There are a few different methods for obtaining oxygen.
-	Electrolysis – relatively high cost, studies show it can be viable with photovoltaic cells if we sell the hydrogen for 10eur/kg, the oxygen is worth 3eur/kg. [https://www.sciencedirect.com/science/article/pii/S0306261921012964]
-	Purchase cylinders @ £41 ex. VAT for 85kg of O2, or 50p/kg [https://www.boconline.co.uk/shop/en/uk/gas-a-z/oxygen/oxygen-cylinder-medical-grade-compressed-gas]
-	PSA/VPSA involves separating air into its constituent components by means of adsorption. In other words, the gas molecules bind with adsorbent material at different rates depending on the pressure. This allows operators to single out one particular gas from air. [https://www.linde-engineering.com/en/images/Oxygen\%20generation.\%20By\%20Vacuum\%20Pressure\%20Swing\%20Adsorption._tcm19-416454.pdf] Refining oxygen from air using vapour pressure swing adsorption has promise for higher consumption tasks such as in hospitals, but studies and simulations show that PSA plants are not viable are not economical at low capacity – quotes estimate prices of $115900-2228000. [https://wobogroup.en.made-in-china.com/product/ldIfpNxPXsYe/China-Best-Price-95-Purity-Oxygen-Vpsa-Plant-Gas-Generation-Equipment.html] [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10993950/] 
Figure 1https://www.wobogroup.com/product/VPSA-Oxygen-GeneratorsO2.html
Air
If using air, other gases and pollutants within the air need to be tracked/filtered, in particular CO2 and carcinogens. This can be handled with filtering and purification and is typically the method that we see in cultured meat scientific papers due to its simplicity and availability. [https://www.nature.com/articles/s41467-023-38593-4]
Gas Sparging
Gas spargers are usually placed at the bottom of the reactor, where small holes emit small bubbles of gas from the supply. The bubble path takes them through the medium, oxygenating the liquid by diffusion with the additional benefit of bubbled mixing.
Bubbled aeration can produce foam on the surface of the culture; thus a foam control system is needed to prevent excessive foam buildup as foam disrupts the proper suspension of microcarriers.
 
Figure 2 Gas sparging bubbling, extract from https://static-content.springer.com/esm/art\%3A10.1038\%2Fs41467-023-38593-4/MediaObjects/41467_2023_38593_MOESM4_ESM.mp4
 
Figure 3https://static-content.springer.com/esm/art\%3A10.1038\%2Fs41467-023-38593-4/MediaObjects/41467_2023_38593_MOESM1_ESM.pdf
 
Headspace Overlay
Some studies use an overlay to pump gas into the headspace of the bioreactor to allow diffusion into the cell media from above. This has been shown to be effective with filtered air as the pumped gas. [https://www.nature.com/articles/s41467-023-38593-4]
Since the headspace overlay method has proven effective, simple and affordable – this is the method we have chosen for our process. We can measure oxygen transfer using Oxygen Quotient [https://www.alicat.com/monitor-outgassed-material-in-bioreactor-headspace/] and use researched methods from the AIChE to enhance oxygen mass transfer [https://aiche.onlinelibrary.wiley.com/doi/epdf/10.1021/bp00015a010]. We’ll use a pump and filter system that shouldn’t cost more than an estimate of [£2000], but it’s important to get a quote should we need a precise number [estimate references?].
 
